Analysis of Violence in Literature
========================================================

Assumptions:
* The data contained in **Emily Project.xlsx** comes from the FrequencyTable
at the top of the sheet
* The primary categories are **Verbal, Rape, PsycMan, Alpha, Beta, Murder,and Other**
* The works are categorized as **chirs, lit, real, dime, and fan**
* The data is summarized as counts of instances in each work
* 0.05 will be used as the alpha value for significance

The data is summarized as follows...

```{r}
require(xlsx)
require(multcomp)

fn <- file.path("c:","users","Rob","downloads","Emily Project.xlsx")
file.exists(fn)

X <- read.xlsx(fn, sheetIndex=2, rowIndex=1:18, colIndex=2:36)

ind <- which(names(X) %in% c("Verbal.Bullying.Verbal.Aggression","Rape",
                             "Psychological.Manipulation",
                             "Alpha.Male.Partner","Beta.Male.Partner","Murder"))
Y <- X[,ind]
names(Y) <- c("Verbal","Rape","PsycMan","Alpha","Beta","Murder")
colSums(Y)
Y$Other <- rowSums(X[,-ind], na.rm=TRUE)

categories <- as.factor(c("chris","chris","lit","lit","real","real","chris","chris","chris",
                        "dime","real","fan","fan","dime","dime","lit","lit"))
by(Y, categories, colSums)
by(Y, categories, sum)
sum(Y)
```

There are multiple possible hypotheses, but only a few should be chosen.

## Test 1
* H0: There is no difference between categories of book based on total counts of violence
* H1: There are differences
* Method: ANOVA
* Result: p > 0.05, therefore the null hypothesis cannot be rejected and there are no
significant differences between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lm1 <- lm(rowSums(X, na.rm=TRUE) ~ categories)
anova(lm1)
```

## Test 2
* H0: There is no difference between categories of book based on counts of violence
in each category of violence
* H1: There are differences
* Method: Multivariate response ANOVA (MANOVA)
* Result: p > 0.05, therefore the null hypothesis cannot be rejected and there are no
significant differences between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
m1 <- manova(as.matrix(Y) ~ categories)
summary(m1)
```

## Test 3
* H0: There is no difference between categories of book based on counts of violence
in one category
* H1: There are differences
* Method: ANOVA

#### Verbal
* Result: p > 0.05, therefore the null hypothesis cannot be rejected and there are no
significant differences between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lmv <- lm(Y$Verbal ~ categories)
anova(lmv)
```
#### Rpe
* Result: p > 0.05, therefore the null hypothesis cannot be rejected and there are no
significant differences between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lmr <- lm(Y$Rape ~ categories)
anova(lmr)
```

#### Psychological Manipulation
* Result: p < 0.05, therefore the null hypothesis can be rejected.  There is at least
one significant difference between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lmp <- lm(Y$PsycMan ~ categories)
anova(lmp)
```

* The significant result is between the lit and chris categories.  The lit category
is siginifantly more violent than the chris category.  Note that the p-values in
these multiple comparisons are adjusted such that the overall error rate is still
0.05.

```{r}
lmp.mc <- glht(lmp, linfct=mcp(categories="Tukey"))
summary(lmp.mc, test=adjusted(type="holm"))
```

#### Alpha Male
* Result: p > 0.05, therefore the null hypothesis cannot be rejected and there are no
significant differences between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lma <- lm(Y$Alpha ~ categories)
anova(lma)
```

#### Beta Male
* Result: p < 0.05, therefore the null hypothesis can be rejected.  There is at least
one significant difference between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lmb <- lm(Y$Beta ~ categories)
anova(lmb)
```

* The significant result is between the lit and fan categories.  The lit category
is siginifantly more violent than the fan category.  Note that the p-values in
these multiple comparisons are adjusted such that the overall error rate is still
0.05.

```{r}
lmb.mc <- glht(lmb, linfct=mcp(categories="Tukey"))
summary(lmb.mc, test=adjusted(type="holm"))
```

#### Murder
* Result: p > 0.05, therefore the null hypothesis cannot be rejected and there are no
significant differences between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lmm <- lm(Y$Murder ~ categories)
anova(lmm)
```

#### Other types of violence
* Result: p > 0.05, therefore the null hypothesis cannot be rejected and there are no
significant differences between categories of book based on counts of violence. (Look at
the Pr(>F) in the table)

```{r}
lmo <- lm(Y$Other ~ categories)
anova(lmo)
```

## Alternate Tests

As a side note, one of the most important assumptions of an ANOVA is that the
data is normally distributed.  This is a difficult assumption to meet when 
the data are counts and they are bounded by zero on the lower side.  Since your
professor suggested an ANOVA, that test is provided.  I would have done the
distribution-free test called Kruskal-Wallis instead.  On the down side, however,
the Kruskal-Wallis test shows no significant differences.

```{r}
kvtotal <- kruskal.test(rowSums(Y, na.rm=TRUE) ~ categories)
kvtotal
kv <- kruskal.test(Y$Verbal ~ categories)
kv
kr <- kruskal.test(Y$Rape ~ categories)
kr
kp <- kruskal.test(Y$PsycMan ~ categories)
kp
ka <- kruskal.test(Y$Alpha ~ categories)
ka
kb <- kruskal.test(Y$Beta ~ categories)
kb
km <- kruskal.test(Y$Murder ~ categories)
km
ko <- kruskal.test(Y$Other ~ categories)
ko
```

