Rational Number Class
========================================================

```{r temp, echo=FALSE, eval=TRUE, results='hide'}
read_chunk(file.path("C:","Users","Rob","Documents","Repositories","RSandbox","rational","rational number class chunks.r"))
```

```{r setupchunk, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
```

This document describes the S4 implementation of a rational number class in R.  
A rational number is a real number which can be expressed as the quotient of two 
integers.  In R, this is expressed as an S4 class with objects:

```{r s4class, echo=TRUE, eval=TRUE}
```
The S4 class, `rational` has slots `numerator, denominator, and value` which
are `integer, integer, and numeric`.  At construction time, the validation
function is called to ensure that the object is created correctly.  When
constructed, the `value` slot should be automatically populated from the 
`numerator` and `denominator` slots, so a custom initialize method is required.

```{r s4classInitialize, echo=TRUE, eval=TRUE, results='hide'}
```

Based on these definitions, a number of exceptions can be thrown depending
on how the object is constructed.

```{r test1, echo=TRUE, eval=TRUE}
```

# Constructor

Next, we would like a generic method that allows the rational number to be
created in a more natural manner, by typing `rational(n,d)` instead of
`new("rational", n, d)`.  Creating a generic method allows one to create
methods that have different signatures.

```{r genericRational, echo=TRUE, eval=TRUE, results='hide'}
```
Now, signatures can be created for the `rational` method that allow for rational
numbers to be created from combinations of `integers` and `numerics`.

```{r methodsRational, echo=TRUE, eval=TRUE, results='hide'}
```
Tests verify the behavior:

```{r test2, echo=TRUE, eval=TRUE}
```
# R Objects
It is natural in R to have a `length` and `[` method for a data type.

```{r length, echo=TRUE, eval=TRUE, results='hide'}
```

```{r test3, echo=TRUE, eval=TRUE}
```
# Accessor Methods

In order to protect the objects as much as possible, it may be necessary to
access the numerator and denominator of an existing object.  Use
* `setNumerator(rational, numeric)`
* `setDenominator(rational, numeric)`
* `getNumerator(rational)`
* `getDenominator(rational)`
* `getValue(rational)`

(Code is hidden due to its simplicity)

```{r accessors, echo=FALSE, eval=TRUE, results='hide'}
```

```{r test4, echo=TRUE, eval=TRUE}
```

# Define S4 Group Generic Functions

## Arith

The `rational` class should define the S4 group generics.  For example, the
`Arith` group contains `r getGroupMembers("Arith")`.  For these methods, either
use the generic method for the arithmetic function with the numeric representation
of the rational number and another numeric argument, or use custom defined
functions for operations with rational numbers.

```{r arith, echo=TRUE, eval=TRUE, results='hide'}
```

```{r loopmethods, echo=FALSE, eval=TRUE, results='hide'}
```

```{r foo, eval=TRUE, echo=TRUE, results='hide'}
````

```{r test5, echo=TRUE, eval=TRUE}
```

## Compare

The compare generics are `r getGroupMembers("Compare")`

```{r compare, echo=TRUE, eval=TRUE, results='hide'}
```

```{r test6, echo=TRUE, eval=TRUE}
```

## Math and Math2

The Math generics are `r getGroupMembers("Math")` and 
`r getGroupMembers("Math2")`.  Most `Math` functions use the `numeric` value,
but some are designed to more accurate with a `rational` number.

```{r math, echo=TRUE, eval=TRUE, results='hide'}
```

```{r test7, echo=TRUE, eval=TRUE}
```

## Summary

The `Summary` generics are `r getGroupMembers("Summary")`

```{r summary, echo=TRUE, eval=TRUE, results='hide'}
```

```{r test8, echo=TRUE, eval=TRUE}
```

## Display

```{r display, echo=TRUE, eval=TRUE, results='hide', message=FALSE}
```

## Conversion

```{r conversion, echo=TRUE, eval=TRUE, results='hide'}
```

```{r test9, echo=TRUE, eval=TRUE}
```
